% =========================================================
% Nearness: Point-Set Topology — Notes Index
% =========================================================

% ---------------------------------------------------------
% Breadcrumb
% ---------------------------------------------------------
\begin{tcolorbox}[
  colback=gray!6,
  colframe=gray!40,
  arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Where You Are in the Journey}},
  fonttitle=\small\bfseries
]
\begin{center}
\small
Propositional Logic
$\;\to\;$ Sets \& Functions
$\;\to\;$ $\mathbb{R}$, Real Analysis
$\;\to\;$ Metric Spaces
$\;\to\;$ \textbf{Nearness (Point-Set Topology)}
$\;\to\;$ Finiteness (Compactness)
$\;\to\;$ $\cdots$
\end{center}

\medskip
\noindent\textbf{How we got here.}
Metric spaces showed us that convergence and continuity can be defined
without coordinates --- only distance matters.
Point-set topology takes the final abstraction step: it discards
distance altogether, retaining only the family of open sets
as the primitive notion.

\medskip
\noindent\textbf{What this chapter builds.}
We begin with the topology inherited from metric spaces --- the content
moved here from the metric-spaces chapter --- and then abstract further.
Topics include the open-set framework, nowhere-dense sets, the Baire
Category Theorem, the Generalized Heine-Borel theorem, and the foundations
of compactness and connectedness.

\medskip
\noindent\textbf{Where this leads.}
Compactness is the global finiteness principle that connects topology
back to analysis: continuous functions on compact sets attain their
extrema and are uniformly continuous, unlocking the Riemann integral.
\end{tcolorbox}
\vspace{1em}

% ---------------------------------------------------------
% Structural Role
% ---------------------------------------------------------
\begin{tcolorbox}[
  colback=gray!6,
  colframe=gray!40,
  arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Structural Role: Nearness --- Neighborhood Structure Without Distance}},
  fonttitle=\small\bfseries
]
Metric spaces taught us that distance controls convergence.
Nearness --- point-set topology --- goes one step further.

It discards the metric entirely
and retains only the family of open sets:
the neighborhoods that determine local behavior.

Convergence, continuity, compactness, and limit points
can all be defined using only this neighborhood structure.

Topology reveals that many sequence-based results
were secretly statements about open sets.
The abstraction completes a three-stage descent:

\begin{itemize}
  \item Order $\;\to\;$ distance,
  \item Distance $\;\to\;$ neighborhoods,
  \item Neighborhoods $\;\to\;$ open sets.
\end{itemize}

Each stage removes a layer of extra structure
while preserving the phenomena that matter for analysis.
Topology isolates the minimal data needed to discuss
convergence and continuity.
\end{tcolorbox}
\vspace{1em}

% ---------------------------------------------------------
% Structural Roadmap
% ---------------------------------------------------------
\subsection*{Structural Roadmap}

\begin{center}
\textbf{Definitions $\longrightarrow$ Main Theorems
$\longrightarrow$ Consequences and Logical Structure}
\end{center}

The development follows the same architecture used throughout this
project, now abstracting neighborhood structure independently of distance.
The global progression is:

\begin{enumerate}
  \item Open sets, closed sets, and their algebra (from metric spaces)
  \item Nowhere-dense sets and the Baire Category Theorem
  \item Generalized Heine-Borel: compact $\iff$ complete $+$ totally bounded
  \item Compactness via open covers
  \item Sequential compactness and the classical Heine-Borel theorem
  \item Connectedness
\end{enumerate}

\begin{remark*}[Primary sources]
Abbott, \textit{Understanding Analysis}, Chapter~3;
Pugh, \textit{Real Mathematical Analysis}, Chapter~2 \S1--6;
Rudin, \textit{Principles of Mathematical Analysis}, Chapter~2.
\end{remark*}

% ---------------------------------------------------------
% Content
% ---------------------------------------------------------
\input{volume-v/topology/point-set-topology/notes/notes-topology}
\input{volume-v/topology/point-set-topology/notes/notes-nowhere-dense-baire}
\input{volume-v/topology/point-set-topology/notes/notes-generalized-heine-borel}

% Future sections (uncomment as content is written):
% \input{volume-v/topology/point-set-topology/notes/notes-compactness}
% \input{volume-v/topology/point-set-topology/notes/notes-connectedness}
